FantasticLamp: quantifying genomic edits with genome variation graphs

### 2024-02-01 paper/main.tex
minor edits

\title[FantasticLamp]{FantasticLamp: a genome graph based pipeline for calculating the efficacy of genomic edits}
\author[Schutte \textit{et~al}.]{

@@

An alternative to these linear methods is the use of a genome graph approach, which can offer a more comprehensive perspective on sequence relationships, avoiding reference bias, and capturing all allele state mixtures, even in cases of overlapping edits \citep{eggertsson2017graphtyper,Martiniano_2020}.
Genome graphs provide a more accurate and complete view compared to linear alignment methods by allowing reads to be mapped to a graph that represents a population of sequences and the variations between them, rather than a single reference \citep{garrison2018variation,paten2017genome}.

@@

The genome graph consists of the reference genome, the intended edit sequences (referred to as `homology arms' or `edit homology arms'), and the reference sequences at the edit sites (termed `reference homology arms').
The core structure of the graph is an alignment of homology and reference homology arms onto the reference genome, which we complet using \textit{minimap2} \citep{li2018minimap2}.
The inclusion of the reference homology arms enables the pipeline to calculate the relative coverage, thereby determining the edit efficiency by comparing the coverage of edit homology arms to that of the reference homology arms.

@@

For further validation, we created a much larger validation test by taking a full yeast genome \citep{ncbis288c2023} and introducing a single edit (ranging from 20 - 25bp) at a specific locus, which we then simulated reads from.
We created 100 of these edited genomes, each with a different edit at a random locus, and simulated 10x coverage, paired-end reads from each genome.
We then ran \textit{FantasticLamp} on two sets of these simulated reads; one set with a base error rate of 0.5\%, and one with a base error rate of 1\%.
In both of these validation tests, \textit{FantasticLamp} was able to identify the correct edit state for 97 out of 100 edits.
The scripts used to generate these genomes and reads (as well as to perform the actual validation tests) are available in the code repository.

>>>>

\title[FantasticLamp]{FantasticLamp: quantifying genomic edits with genome variation graphs}
\author[Schutte \textit{et~al}.]{

@@

An alternative to these linear methods is the use of a genome graph approach, which can offer a more comprehensive perspective on sequence relationships, avoiding reference bias, and capturing all allele state mixtures, even in cases of overlapping edits \citep{eggertsson2017graphtyper,Martiniano_2020}.
Genome graphs provide a more accurate and complete view compared to linear alignment methods by allowing reads to be mapped to a graph that represents a population of sequences and the variations between them, rather than a single reference \citep{garrison2018variation,paten2017genome,Eizenga_2020}.

@@

The genome graph consists of the reference genome, the intended edit sequences (referred to as ``homology arms'' or ``edit homology arms''), and the reference sequences at the edit sites (termed ``reference homology arms'').
The core structure of the graph is an alignment of homology and reference homology arms onto the reference genome, which we complete using \textit{minimap2} \citep{li2018minimap2}.
The inclusion of the reference homology arms enables the pipeline to calculate the relative coverage, thereby determining the edit efficiency by comparing the coverage of edit homology arms to that of the reference homology arms.

@@

For further validation, we created a much larger validation test by taking a full yeast genome (GCF\_000146045.2) and introducing a single edit (ranging from 20 - 25bp) at a specific locus, which we then simulated reads from.
We created 100 of these edited genomes, each with a different edit at a random locus, and simulated 10x coverage, paired-end reads from each genome.
We then ran \textit{FantasticLamp} on two sets of these simulated reads; one set with a base error rate of 0.5\%, and one with a base error rate of 1\%.
In both of these validation tests, \textit{FantasticLamp} was able to identify the correct edit state for 97 out of 100 edits.
The scripts used to generate these genomes and reads (as well as to perform the actual validation tests) are available in the code repository.
